Leaf traits drive differences in above-ground biomass distribution among major plant functional types

### 2015-11-11 manuscript.Rnw
almost there

\section{New list of figures}
\begin{enumerate}
\item Tree diagram plus climate space of dataset. Unchanged from before, except remove MT (?).
\item MF and MS as function of tree height. Reviewer 2 asks "But are differences in Figure 2 statistically signficant?", weird question and subsequent comments. Figure could add trend lines (as smoothers), but refer to later figures for significant differences. The comment that sampling may have been uneven is correct, which is why we correct for plant height in the analyses.
\item (a) MF/MS vs. plant height by PFT (was LMF) (b) and (c) as before, except for MF/MS. Reviewer 2 wants us to remove panels d and e, which we can do because we won't need MF/AS as a proxy anymore (should be put in SuppInfo instead).
\item Histograms by PFT. This was nice when we used size-invariant MF/AS, but we can still show LMA and AF/AS? The first has large PFT differences, the second not any.
\item Not sure how current Fig 5 will fit in the new version
\item Not sure how current Fig 6 will fit in the new version, since we won't be able to use MF/AS. Can make this plot for AF/AS though, since reviewer 1 thinks that has biological relevance (and MF/AS doesn't, for some reason).
\end{enumerate}

@@

% 200 word max.
\begin{enumerate}
\item The partitioning of biomass into leaves and stems is an important component of  global vegetation models (GVMs). Although GVMs typically assume that the major woody plant functional types (PFTs) differ in biomass partitioning, empirical studies have not been able to justify these differences. Here we test for differences between PFTs in distribution of biomass between leaves and stems.
\item We use the recently published Biomass And Allometry Database (BAAD), a large database including observations for individual plants. The database covers the global climate space, allowing us to test for direct climate effects in addition to PFT. 
\item The leaf mass fraction (LMF, leaf / total aboveground biomass) varied strongly between PFTs (as defined by deciduous vs. evergreen and gymnosperm vs. angiosperm). We found that LMF, once corrected for plant height, was proportional to leaf mass per area across PFTs. As a result, the PFTs did not differ in the amount of leaf area supported per unit above ground biomass. We found only weak and inconsistent effects of climate on biomass partitioning.
\item Combined, these results will allow for systematic benchmarking of biomass partitioning routines in GVMs.
\end{enumerate}

\emph{Keywords}: allocation, plant allometry, biomass estimation, specific leaf area, dynamic global vegetation model

@@

The partitioning of forest biomass among leaves and stems strongly influences the productivity and carbon cycle of the world's vegetation \citep{ise2010, dekauwe2014,friend2014}. Biomass stored in woody stems has a long residence time \citep{luyssaert2008}, whereas leaf biomass turns over quickly, entering the soil carbon cycle where the majority of carbon is released back to the atmosphere \citep{ryanlaw2005}. Globally, forests store approximately 360Pg of carbon in living biomass \citep{panetal2011}, equivalent to almost 40 years of current anthropogenic CO\textsubscript{2} emissions \citep{friedlingsteinetal2014}. Reducing uncertainties about biomass partitioning and carbon residence times in GVMs is therefore a priority for understanding the effects of climate and other environmental change on the global carbon cycle \citep{friend2014,dekauwe2014,negron2015}.

% less focus on GVMs; focus on ecological/biological reasons to test these things
% ... but make sense to look at PFTs though, and this is the actual reason for doing that?
Perhaps the biggest challenge for GVMs is to capture the combined responses of the more than 250,000 plant species comprising the world's vegetation. While most plants have the same basic resource requirements and physiological function, large differences exist among species in the amount of energy invested in different tissues (leaves, stems, roots). The approach taken by most GVMs for dealing with this functional diversity is to consider only a few archetypal plant functional types (PFTs) \citep{harrisonetal2010,wullschleger2014}, assumed to differ in key physiological attributes. While GVMs assume or predict differences in biomass partitioning between PFTs (Notes S1), these differences are poorly constrained, due to limited available data. Moreover, there is little consensus on how biomass partitioning and allocation (see Methods for terminology) should be modelled in GVMs \citep{franklin2012,dekauwe2014,friend2014}. These shortcomings largely reflect the lack of suitable datasets of global scope with which models can be tested, constrained and compared \citep{wolf2011}.

%explain why stem/leaf biomass distribution is important here.
In this paper we are primarily interested in the distribution of biomass between leaves and woody stems, an important component of the residence time of carbon in ecosystems \citep{friend2014}. Previous work based on either whole stands \citep{oneilldeangelis1981,enquistniklas2002,reich2014} and or a mix on stand-- and individual-based measurements \citep{poorter2012, poorter2015}, reported differences between angiosperms and gymnosperms in the amount of leaf biomass per unit above-ground biomass (the 'leaf mass fraction', LMF). \citet{poorter2012} and \citet{enquistniklas2002} attributed higher LMF in gymnosperms to longer leaf lifespan (LL) compared to typical angiosperms, but this begs the question whether differences in LMF are also apparent between deciduous and evergreen functional types within angiosperms. It is also unknown whether PFTs with higher leaf biomass also have higher total leaf area, which is relevant because leaf area drives total light interception and thus productivity. Some oft-cited studies have also assumed that gymnosperms carry more leaf area than angiosperms \citep{chabothicks1982, bond1989}.

Little is known about global-scale patterns in LMF and LAR in relation to climate. It can be expected that biomass partitioning is correlated with precipitation or mean annual temperature because smaller-scale comparisons have shown responsiveness of biomass partitioning to environmental drivers \citep{berninger1994,callaway1994,delucia2000,poyatos2007}. Moreover, a recent study by \citet{reich2014} demonstrated that stand-scale biomass partitioning was related to mean annual temperature across diverse forest stands. Indeed, many GVMs assume that climate affects biomass partitioning within a given PFT, often through soil water stress or other abiotic stress factors (see Notes S1). Again, however, these models do not agree on the degree of plasticity in biomass partitioning, or which climate variables it should respond to.

% integrate poorter2015 and reich2014 better
Despite considerable advances in theory underlying allometric scaling in plants \citep{enquist2003, west1999, savage2010} we do not yet have a clear understanding of potential differences between PFTs in terms of biomass partitioning. Instead, most previous work has focussed on understanding size-related shifts in biomass partitioning, as governed by constraints including hydraulic supply and mechanical stability \citep{savage2010}. Rather than to advance a specific model of biomass partitioning or allocation, our view is that a broader evidence base is needed first to elucidate patterns between PFTs. This should enable those building GVMs to refine algorithms and parameter values to more closely match the observations. Here, we use a unique, new database \citep{falster2015} (Fig.~\ref{fig:figoverview}) to establish general rules on how biomass partitioning differs among three dominant woody PFTs across the globe: evergreen gymnosperms, evergreen angiosperms, and deciduous angiosperms. A recent compilation of plant biomass data \citep{poorter2015} speculated that differences between gymnosperms and angiosperms in distribution of biomass between leaves and stems is related to differences in leaf mass per area. Here we can directly test this hypothesis as our database, unlike those of \citet{enquistniklas2002}, \citet{reich2014}, and \cite{poorter2015}, includes many observations of leaf area as well as leaf mass measured on the same plants.

A second objective was to study relationships between biomass partitioning and climate variables or biome at a global scale. The dataset includes observations of biomass and size metrics for individual plants, compiled from 175 studies across nine vegetation types (Fig. S1), across the three major biomes (boreal, temperate and tropical). In this paper we focus on field-grown woody plants, spanning the entire size range of woody plants (0.01 - \textgreater 100 m height). One key challenge is thus to account for the very large size variation commonly found in any allometric variable \citep{niklas1994}. We do this by fitting a semi-parametric data-driven statistical model, i.e. one that does not assume a particular functional form, thereby allowing us to study PFT patterns at a common plant height.

@@

% no longer needed? we will use biomass distribution instead of partitioning to be consistent with Poorter 2015.
\subsection*{Terminology}
\label{sec:partterm}

The terms 'partitioning' and 'allocation' have been used in various ways, confusing comparisons between studies \citep{litton2007}. Here, we define biomass partitioning as the actual distribution of biomass between compartments (e.g., leaves vs. stems), and biomass allocation as the proportion of net primary production (NPP) that is allocated to some compartment. The two processes are different because of continuous turnover of biomass, which differs strongly between compartments. We may write \citep{mcmurtriewolf1983},

$$
\frac{dM_i}{dt} = a_i \textrm{NPP} - \tau_i M_i
$$

where $M_i$ is the biomass in some compartment (leaves, stems or roots) remaining on the plant, $a_i$ the annual allocation of NPP to that compartment, and $\tau_i$ the annual turnover (or loss) of compartment $i$ from the plant. It is thus easy to see that partitioning ($M_i/M_T$, where $M_T$ is total biomass) can be different from allocation ($a_i$) because turnover ($\tau_i$) differs between leaf and wood biomass. Here, we present data on biomass partitioning, which can inform models of allocation only when estimates of turnover are available.

@@

For the above-ground biomass pools, we calculated the leaf mass fraction ($M_F / M_T$, where $M_T$ is total above-ground biomass) and leaf area ratio ($A_F / M_T$). These variables are related by,

$$\textrm{LMF} = \frac{M_F}{M_T} = \left(\frac{M_F}{A_F}\right) \left(\frac{A_F}{M_T}\right) = \textrm{LMA} \times \textrm{LAR}.$$

@@

Despite the exceptional size of our dataset, the strong size-dependence of LMF still hinders comparisons across climatic gradients, due to the small sample size available within each species or site when sampling at a common height. To study climate effects on biomass partitioning, we therefore further decomposed LMF as the product,

\begin{equation}\label{eq:M1}
\textrm{LMF} = \left(\frac{M_F}{A_S}\right) \left(\frac{A_S}{M_T}\right).
\end{equation}

This decomposition showed that the vast majority of size-related variation was captured by $A_S/M_T$ alone, indicating that comparisons among $M_F/A_S$ could be made without needing to compare at a common height. A similar decomposition  of LAR as the product of $A_F/A_S$ and $A_S/M_T$ produced the same outcome. We therefore analysed for climatic effects on $M_F/A_S$ and $A_F/A_S$ in two ways: with quantitative variables (mean annual precipitation, MAP; mean annual temperature, MAT), and by biome (tropical / temperate/ boreal), a simple classification taking into account both MAT and MAP (Fig. S1). In both cases we used PFT and plant height as covariates varying across climate space.

@@

The raw data in Fig.~\ref{fig:figrawdata} show a steeper increase of woody aboveground biomass ($M_S$) with plant height compared to foliage biomass ($M_F$). As a result, LMF decreased with plant height, with the three plant functional types clearly differing in LMF across nearly the entire size range (Fig.~\ref{fig:figlmf}a). When we corrected for plant height by estimating LMF at a common plant height, we found that LMF was proportional to the average leaf mass per area (LMA) across the three PFTs (Fig.~\ref{fig:figlmf}b). As a consequence, LAR was invariant between PFTs, because LAR = LMF / LMA (Fig. ~\ref{fig:figlmf}c and Fig. S2).

Mirroring the results for LMF, we found that the amount of leaf mass per unit stem area ($M_F/A_S$) differed among the three PFTs, that these differences were also proportional to LMA (Fig. ~\ref{fig:figlmf}d), and that PFTs were similarly invariant in the amount of leaf area per unit stem area ($A_F/A_S$) (Fig.~\ref{fig:figlmf}e).

The above results highlight how LMA drives differences in biomass partitioning among dominant woody PFTs. To assess whether LMA could replace PFT when predicting biomass partitioning, we fit a linear mixed-effects model to LMF and $M_F/A_S$ using PFT and plant height (and the quadratic term) as predictors (and all interactions). We then replaced PFT with LMA, and found that the model with LMA could explain almost as much variation in LMF as the model with PFT ($R^2$ = 0.74 with PFT vs. 0.62 with LMA), likewise for $M_F/A_S$ ($R^2$ = 0.29 vs. 0.28).

There was considerable variation in all studied variables between species within PFTs (Fig. ~\ref{fig:fighist}). To understand whether we can improve on a PFT-based classification by including other traits that affect biomass partitioning, we decomposed $M_F/A_S$ into LMA ($M_F/A_F$), and the ratio of leaf area to stem cross-sectional area ($A_F/A_S$):

\begin{equation}\label{eq:M2}
\frac{M_F}{A_S} = \left(\frac{M_F}{A_F}\right) \left(\frac{A_F}{A_S}\right).
\end{equation}

The ratio $M_F/A_S$ is relevant because plant leaf mass is frequently estimated from $A_S$ with allometric equations \citep{shinozaki1964b, chave2005}, using records of stem diameter commonly recorded on long-term monitoring plots.  The correlation between $M_F/A_S$ and LMA was found to hold also within PFTs (Fig.~\ref{fig:figlmfspecies}a), explaining 30\% of the variation in $M_F/A_S$ across species. The regressions across species by PFT were broad and overlapping, with a more general relationship extending across the entire LMA axis. A larger fraction of the variation in $M_F/A_S$ was explained by species-level differences in $A_F/A_S$ (Fig.~\ref{fig:figlmfspecies}b). As previously noted, $A_F/A_S$ does not differ systematically among PFTs (Fig.~\ref{fig:figlmf}e), but does vary close to two orders of magnitude across species. In Fig.~\ref{fig:figlmfspecies}b, the separation among PFTs in $M_F/A_S$ arises due to differences in LMA.

We found that decomposing LMF and LAR as shown in Eq. \ref{eq:M1} was very useful to study climate effects because the second term ($A_S/M_T$) absorbed nearly all of the size-related variation in LMF (Table \ref{table:statstable} and Fig. S3) and is otherwise fairly conserved across PFTs (Fig. S3). As shown in Fig.~\ref{fig:figlmfspecies}b and Fig.~\ref{fig:figlmf}d, the term $M_F/A_S$ exhibits comparable differences between PFTs to LMF, but unlike LMF, $M_F/A_S$ is nearly independent of plant height (Table \ref{table:statstable}). $M_F/A_S$ is thus a useful proxy for LMF that can be compared across species and sites.

Regardless of how we analysed the data, we found only weak climatic effects on biomass partitioning within PFTs (Fig.~\ref{fig:figclimate}). Biome consistently explained very little variation when added to a statistical model in addition to PFT and plant height ($R^2$ increased by only 0.01 - 0.06, see Table \ref{table:statstable}). Likewise, we found weak and inconsistent effects of climate variables (MAP and MAT) (Fig.~\ref{fig:figclimate}b-c). These variables explained little variation when added to a statistical model in addition to PFT and plant height ($R^2$ increased by 0.01 and 0.11, respectively, see Table \ref{table:statstable}).

@@

\textbf{Fig. S1} Global coverage of the climate space by the dataset, labelled by vegetation type.\\
\textbf{Fig. S2} Leaf area ratio ($A_F/M_T$) by PFT.\\
\textbf{Fig. S3} Relationship between above-ground biomass and basal stem area.\\
\textbf{Notes S1} Modelling of biomass partitioning in global vegetation models (GVMs)

>>>>

\section{New list of figures}
\begin{enumerate}
\item Tree diagram plus climate space of dataset. Unchanged from before, except remove MT (?).
\item MF and MS as function of tree height. Reviewer 2 asks "But are differences in Figure 2 statistically signficant?", weird question and subsequent comments. Figure could add trend lines (as smoothers), but refer to later figures for significant differences. The comment that sampling may have been uneven is correct, which is why we correct for plant height in the analyses.
\item (a) MF/MS vs. plant height by PFT (was LMF) (b) and (c) as before, except for MF/MS. Reviewer 2 wants us to remove panels d and e, which we can do because we won't need MF/AS as a proxy anymore (should be put in SuppInfo instead).
\item AF vs. AS by pft, for all data, with fitted SMA. This shows AF is proportional to AS (approx, except larger plants), and no/very very small effect of PFT.
\item Stem mass as function of AS, PFT and height. a) Raw data (to link to main equation) show no clear PFT effect. b) MS / AS is nearly proportional to H, as expected from simple stem mass equation. c) for larger plants (H*AS > 0.01m3), there is a significant PFT effect, with EG having lower MS / (AS*H) (see d) for averages across entire subset shown in c). This can be explained from differences in wood density; gymnosperms have lower wood density on average.
\item Histograms of the three main variables: AF/AS, MF/AF, and height-corrected stem mass index, MS/(AS*H).
\end{enumerate}

@@

Much too long, of course!
% 200 word max.
\begin{enumerate}
\item The partitioning of biomass into leaves and stems is an important component of  global vegetation models (GVMs).
\item We use the recently published Biomass And Allometry Database (BAAD), a large database including observations for individual plants. The database covers the global climate space, allowing us to test for direct climate effects in addition to PFT. 
\item The ratio of leaf mass to woody above-ground mass (MF/AS) varied strongly between PFTs (as defined by deciduous vs. evergreen and gymnosperm vs. angiosperm). We found that MF/MS, once corrected for plant height, was proportional to leaf mass per area across PFTs. As a result, the PFTs did not differ in the amount of leaf area supported per unit above ground biomass. 
\item We partitioned the variance of MF/MS into three component variables (AF/AS, MF/AF and MS/AS), and each of these four variables into PFT, climate and plant height. Plant height explained most variation in MF/MS and MS/AS (but not H or MF/AF). PFT explained much variation in MF/MS and MF/AF (but not AF/AS and only for larger plants in MS/AS). 
\item Climate variables (MAT, MAP and its interaction) consistently explained very little additional variation, demonstrating very weak control of climate on biomass distribution and leaf area. We argue that climate effects will largely be indirectly through plant height.
\item Combined, these results show consistent patterns in above-ground biomass distribution and leaf area relationships among major woody plant functional types, and can be used to further constrain GVMs.
\end{enumerate}

\emph{Keywords}: allocation, plant allometry, biomass estimation, specific leaf area

@@

\begin{itemize}
\item Biomass allocation is important; need to improve GVMs and make them more consistently benchmarked against global observations. 
\item Plant functional types; can expect differences in biomass partitioning and leaf area relationships. Useful from viewpoint of ecology (functional consequences of angiosperm vs. gymnosperm divide; evergreen vs. deciduous?). Also useful for GVMs which use PFT for parameterization.
\item Enquist&Niklas showed differences in leaf mass at given stem mass between gymnosperms and angiosperms.Poorter2015 showed X. Reich2014 showed Y. Open questions are 1,2,3. One major gap is that these previous studies did not include estimates of whole-plant leaf area. Leaf area is relevant because bla, but may follow very different patterns from leaf mass. Poorter2015 speculated that X.
\item What about climate? We know from small-scale experiments and observational studies that biomass distribution varies with climate drivers. But unclear if global patterns persist. Reich2014 showed X, consistent with glasshouse stuff from Poorter2015. 
\item We have the following objectives:
\begin{enumerate}
\item Test whether MF/MS is proportional to leaf mass per area (as suggested by Poorter)
\item Partition the variance of leaf/stem biomass distribution into component variables, and into plant height (which is important, obviously!), plant functional type and climate. 
\end{enumerate}

\end{itemize}

% The partitioning of forest biomass among leaves and stems strongly influences the productivity and carbon cycle of the world's vegetation \citep{ise2010, dekauwe2014,friend2014}. Biomass stored in woody stems has a long residence time \citep{luyssaert2008}, whereas leaf biomass turns over quickly, entering the soil carbon cycle where the majority of carbon is released back to the atmosphere \citep{ryanlaw2005}. Globally, forests store approximately 360Pg of carbon in living biomass \citep{panetal2011}, equivalent to almost 40 years of current anthropogenic CO\textsubscript{2} emissions \citep{friedlingsteinetal2014}. Reducing uncertainties about biomass partitioning and carbon residence times in GVMs is therefore a priority for understanding the effects of climate and other environmental change on the global carbon cycle \citep{friend2014,dekauwe2014,negron2015}.
% 
% % less focus on GVMs; focus on ecological/biological reasons to test these things
% % ... but make sense to look at PFTs though, and this is the actual reason for doing that?
% Perhaps the biggest challenge for GVMs is to capture the combined responses of the more than 250,000 plant species comprising the world's vegetation. While most plants have the same basic resource requirements and physiological function, large differences exist among species in the amount of energy invested in different tissues (leaves, stems, roots). The approach taken by most GVMs for dealing with this functional diversity is to consider only a few archetypal plant functional types (PFTs) \citep{harrisonetal2010,wullschleger2014}, assumed to differ in key physiological attributes. While GVMs assume or predict differences in biomass partitioning between PFTs (Notes S1), these differences are poorly constrained, due to limited available data. Moreover, there is little consensus on how biomass partitioning and allocation (see Methods for terminology) should be modelled in GVMs \citep{franklin2012,dekauwe2014,friend2014}. These shortcomings largely reflect the lack of suitable datasets of global scope with which models can be tested, constrained and compared \citep{wolf2011}.
% 
% %explain why stem/leaf biomass distribution is important here.
% In this paper we are primarily interested in the distribution of biomass between leaves and woody stems, an important component of the residence time of carbon in ecosystems \citep{friend2014}. Previous work based on either whole stands \citep{oneilldeangelis1981,enquistniklas2002,reich2014} and or a mix on stand-- and individual-based measurements \citep{poorter2012, poorter2015}, reported differences between angiosperms and gymnosperms in the amount of leaf biomass per unit above-ground biomass (the 'leaf mass fraction', LMF). \citet{poorter2012} and \citet{enquistniklas2002} attributed higher LMF in gymnosperms to longer leaf lifespan (LL) compared to typical angiosperms, but this begs the question whether differences in LMF are also apparent between deciduous and evergreen functional types within angiosperms. It is also unknown whether PFTs with higher leaf biomass also have higher total leaf area, which is relevant because leaf area drives total light interception and thus productivity. Some oft-cited studies have also assumed that gymnosperms carry more leaf area than angiosperms \citep{chabothicks1982, bond1989}.
% 
% Little is known about global-scale patterns in LMF and LAR in relation to climate. It can be expected that biomass partitioning is correlated with precipitation or mean annual temperature because smaller-scale comparisons have shown responsiveness of biomass partitioning to environmental drivers \citep{berninger1994,callaway1994,delucia2000,poyatos2007}. Moreover, a recent study by \citet{reich2014} demonstrated that stand-scale biomass partitioning was related to mean annual temperature across diverse forest stands. Indeed, many GVMs assume that climate affects biomass partitioning within a given PFT, often through soil water stress or other abiotic stress factors (see Notes S1). Again, however, these models do not agree on the degree of plasticity in biomass partitioning, or which climate variables it should respond to. 
% 
% % integrate poorter2015 and reich2014 better
% Despite considerable advances in theory underlying allometric scaling in plants \citep{enquist2003, west1999, savage2010} we do not yet have a clear understanding of potential differences between PFTs in terms of biomass partitioning. Instead, most previous work has focussed on understanding size-related shifts in biomass partitioning, as governed by constraints including hydraulic supply and mechanical stability \citep{savage2010}. Rather than to advance a specific model of biomass partitioning or allocation, our view is that a broader evidence base is needed first to elucidate patterns between PFTs. This should enable those building GVMs to refine algorithms and parameter values to more closely match the observations. Here, we use a unique, new database \citep{falster2015} (Fig.~\ref{fig:figoverview}) to establish general rules on how biomass partitioning differs among three dominant woody PFTs across the globe: evergreen gymnosperms, evergreen angiosperms, and deciduous angiosperms. A recent compilation of plant biomass data \citep{poorter2015} speculated that differences between gymnosperms and angiosperms in distribution of biomass between leaves and stems is related to differences in leaf mass per area. Here we can directly test this hypothesis as our database, unlike those of \citet{enquistniklas2002}, \citet{reich2014}, and \cite{poorter2015}, includes many observations of leaf area as well as leaf mass measured on the same plants.
% 
% A second objective was to study relationships between biomass partitioning and climate variables or biome at a global scale. The dataset includes observations of biomass and size metrics for individual plants, compiled from 175 studies across nine vegetation types (Fig. S1), across the three major biomes (boreal, temperate and tropical). In this paper we focus on field-grown woody plants, spanning the entire size range of woody plants (0.01 - \textgreater 100 m height). One key challenge is thus to account for the very large size variation commonly found in any allometric variable \citep{niklas1994}. We do this by fitting a semi-parametric data-driven statistical model, i.e. one that does not assume a particular functional form, thereby allowing us to study PFT patterns at a common plant height. 
%

@@

@@

@@

@@

\begin{itemize}
\item Dataset covers climate space pretty well. 
\item MF and MS are non-linearly related to plant height (Fig 2), as expected. Also note stem mass increases with height more quickly than leaf mass.
\item Leaf/stem biomass distribution was very different between PFTs (Fig 3a), and proportional to LMA when estimated at a common height, so that leaf area ratio did not differ between PFTs. Consistent with variance partitioning: MF/MS mostly explained by H and PFT, but LAR not much explained by PFT.
\item 
\end{itemize}

@@

% \textbf{Fig. S1} Global coverage of the climate space by the dataset, labelled by vegetation type.\\
% \textbf{Fig. S2} Leaf area ratio ($A_F/M_T$) by PFT.\\
% \textbf{Fig. S3} Relationship between above-ground biomass and basal stem area.\\
% \textbf{Notes S1} Modelling of biomass partitioning in global vegetation models (GVMs)
